Make any stupid plot you want

### 2021-05-09 draft/sections/math.tex
clean ups

\section{Topological Equivariant Artist Model}
\label{sec:math}
To guide the implementation of structure preserving visualization components, we develop a mathematical formalism of visualization that specifies how these components preserve \textit{continuity} and \textit{equivariance}. Inspired by the analogous classes in Matplotlib\cite{hunterArchitectureOpenSource}, we call the transformation from data space to graphic space that these building block components implement the \textit{artist}.
\begin{equation}
    \label{eq:artist}
    \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
\end{equation}
The \textit{artist} $\mathscr{\vartist}$ is a map from data $\mathscr{\dtotal}$ to graphic $\mathscr{\gtotal}$ fiber bundles. To explain how the \textit{artist} is a structure preserving map from data to graphic, we first model data (\autoref{sec:math:data}) and graphics (\autoref{sec:math:graphic}) as topological structures that encapsulate component types and continuity. We then discuss the funtional maps from graphic to data (\autoref{sec:math:graphic:base}), data components to visual components (\autoref{sec:math:artist:nu}), and visual components into graphic (\autoref{sec:math:artist:q}) that make up the artist.

@@

A monoid \cite{Monoid2021} $\monoid$ is a set with a binary operation $\ast:\monoid \times \monoid\rightarrow \monoid$ that satisfies the axioms:
\begin{align*}
    \textbf{associativity}\; & \text{for all} a, b, c \in \monoid \, (a \bullet b) \bullet c = a \bullet (b \bullet c)\\
    \textbf{identity}\; & \text{for all } a\in \monoid,\,  e\bullet a = a 
\end{align*} 
As defined on a component of \dfiber, a left monoid action \cite{SemigroupAction2021,nlab:action} of $\monoid_i$ is a set $\dfiber_i$ with an action $\bullet: \monoid\times \dfiber_i \rightarrow \dfiber_i$ with the properties:
\begin{align*}
    \textbf{associativity}\;& \text{for all } f,g \in \monoid_i \text{ and } x\in \dfiber_i,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
    \textbf{identity}\;& \text{for all } x\in \dfiber_i, e\in \monoid_i,\,  e\bullet x = x 
\end{align*}
We conjecture that given these properties, if there is a partial ordering on both sides of the action then the action is equivariant and monotonic.

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/monoid_total.png}
        \caption{}
        \label{fig:math:data:monoid:total}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=1\textwidth]{figures/math/monoid_partial.png}
        \caption{}
        \label{fig:math:data:monoid:partial}
    \end{subfigure}
    \caption{The action $\bullet$ in \autoref{fig:math:data:monoid:total} is the arrows from the partial order diagram of weather states on the left to the diagram of weather states on the right. The action maps the weather states to themselves, thereby preserving the ordering defined by the monoid $\ast$ on both sides of the action. We conjecture that the action in \autoref{fig:math:data:monoid:partial} is monotonic when the partial ordering of the weather states is the same as the partial order of the elements they are mapped to. Given $sun \geq clouds\geq rain,\, lightining$ on the right, we conjecture that the action  $sun,\, clouds \rightarrow clouds$, and $rain,\, lightining \rightarrow rain$ is structure preserving such that the relative ordering of elements is the same as the elements they are mapped to.}
    \label{fig:math:data:monoid}
\end{figure}

@@

As with the fiber \dfiber\, the total monoid space \monoid\ is the cartesian product
\begin{equation}
    \label{eq:math:data:monoid:decompose}
\monoid= \monoid_{0} \times \ldots \times \monoid_{i}\times \ldots \times\ldots \monoid_{n}
\end{equation}
of each monoid $\monoid_{i}$ on $\dfiber_{i}$.  The monoid is added to the specification of the fiber $(\fname_i,\, \ftype_i,\, \fttype\, \monoid_i)$

\subsubsection{Continuity of the Data $K$} 
\label{sec:math:data:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space \dbase\ encodes the continuity of the data space. The table of discrete weather station records has discrete continuity such that each record maps to a single point (a). A gaussian has a value at all points along the interval x is sampled from and therefore has a 1D continuity (b). The globe has a value at all points (latitude, longitude) on the globe and therefore has 2D continuity (c). }
    \label{fig:math:data:base:types}
\end{figure}

The base space \dbase\ provides a way to explicitly encode the continuity of the data, as described in \autoref{sec:intro:data:continuity}. This explicit topology is a concise way of distinguishing between visualizations that appear identical but assume different continuity, for example heatmaps and images. The base space \dbase\ acts as an indexing space, as emphasized by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, \dbase\ can have any number of dimensions and can be continuous or discrete. Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $\dbasepoint \in \dbase$ has a corresponding fiber $\dfiber_k$\cite{QuotientSpaceTopology2020}.

@@

In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping $\pi: \dtotal\rightarrow \dbase$ is the closed interval $\left[0,1\right]$. While the total space \dtotal\ may have components in \dfiber\ that describe any given point $\dbasepoint \in \dbase$, such as \textit{time, latitutde, longitude}, these labels are indexed into from \dbase\ the same as any other components. In contrast to the structural \textit{keys} with associated \textit{values} proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, our model treats keys \dbasepoint\ as a pure reference to topology. Decoupling the keys from their semantics allows the components identifying the keys to be altered, which provides for coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example total rainfall is dependent on time of day and how much rain has already fallen, and changing the coordinate system or time resolution should have no effect on how the records are connected to each other.

@@

\subsubsection{Data \dsection}
\label{sec:math:data:section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section can be decomposed as 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber (\autoref{eq:math:data:base:decompose}, \autoref{eq:math:data:fiber:decompose}), we can decompose \dsection\ as 
\begin{equation}
\label{eq:math:data:section:decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \caption{ Fiber (time, precipitation) with a 1D continuous \dbase\  defined on an interval $\left[0,n\right]$. The sections $\dsection^{(1)}$ and $\dsection^{(2)}$ are constrained such that the time variable must be monotonic, which means each section is a timeseries of precipitation values. They are included in the global set of sections  $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$}
    \label{fig:math:data:sections}
\end{figure}

In \autoref{fig:math:data:sections}, the fiber is the same encoding of $(time, \, precipitation)$ illustrated in \autoref{fig:math:data:fiber}, and the base space is the interval \dbase\ shown in \autoref{fig:math:data:base}. The section $\dsection^{(1)}$ is a function that for a point \dbasepoint\ returns a record in the fiber \dfiber. The section applied to a set of points in \dbase\ resolves to a series of monotonically increasing in time records of \textit{(time, precipitation)} values. Section $\dsection^{(2)}$ returns a different timeseries of \textit{(time, precipitation)} values. Both sections are included in the global set of sections $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$.

@@

\subsubsection{Applications}
Using fiber bundles as our data abstraction allows the model to be applicable to a wide variety of fields and use cases. For example, the section can be any instance of a numpy array\cite{harris2020array} that stores an image, such as an image where the \dbase\ is a 2D continuous plane and the \dfiber\ is $(\reals^{3}, \reals, \reals)$. In this fiber, the $\reals^3$ components encode color, and the other two components are the x and y positions of the sampled data in the image. The continuity of the image is implicitly encoded in the array as the index, so the position components encode the resolution. Instead of an image, the numpy array could also store a 2D discrete table. The fiber may not change, but the \dbase\ would now be 0D discrete points. These different choices in topology indicate, for example, what sorts of interpolation would be appropriate when visualizing the data. Labeled containers can also be described in this framework because of the schema like structure of the fiber. One such example is a pandas series which stores a labeled list, another is a  dataframe\cite{jeff_reback_2020_3715232} which has the structure of a relational table. A series could store the values of $\dsection^{(1)}$ and a second series could be  $\dsection^{(2)}$, while a dataframe would have multiple components and each data frame would be a unique section \dsection. The ability to encode complexity in continuity and components is particulary beneficial when working with  N dimensional labeled data containers. For example, an xarray\cite{hoyer2017xarray} data cube that stores precipitation would be a section of a fiber bundle with a \dbase\ that is a continuous volume and components \textit{(time, latitude, longitude, precipitation)}. This section does not need to resolve to values immediately and instead can be an instance of a distributed data container, such as a dask array \cite{rocklinDaskParallelComputation2015}. Our model provides a common formalism for describing widely used data containers without sacrificing the semantic structure embedded in each container.

@@

\subsubsection{Graphic \gsection}
\label{sec:math:graphic:section}
The section $\gsection: \gbase \rightarrow \gtotal$ is the graphic in an idealized prerender space and also acts as a specification for rendering the graphic to target display format. To demonstrate the role of \gsection\, it is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{x,y}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. The color is the result of the integration over  $\gbase_p$.

@@

\subsubsection{Visual Encoders \vchannel}
\label{sec:math:artist:nu}
We propose that the map from data components to graphic components $\vchannel: \dsection \mapsto \vsection$ is a monoid \textit{equivariant} map. By specifying this constraints, we can guarantee that the stage of the artist that transforms data components into graphic representations is equivariant. These constraints then guide the implementation of reusable component transformers \vchannel\ that are composed when generating the graphic. We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. Therefore, a validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
such that applying equivariant monoid actions to $\dtotal_{i}$ and $\vtotal_{i}$ preserves the map $\vchannel_{i}: \dtotal_{i}\rightarrow \vtotal_{i}$. In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\label{eq:math:artist:nu:equivariance}
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, 
without a loss of generality we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function. We can make this assumption because we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can then compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$, which lets us incorporate $\varphi$ into the action $\bullet$ such that $\varphi$ does not need to be explicitly defined in the constraints.

@@

\subsubsection{Visualization Assembly}
Having described the maps to components in \autoref{sec:math:artist:nu}, we now specify the assembly function \vmarkd\ that composites components in \vtotal\ into a graphic in \gtotal. Since the component transformes \vchannel\ are equivariant, the equivariance constraints carry through to \vmarkd. We specify these constraints to guide the implementation of library components responsible for generating graphics.

@@

\subsubsection{Assembly \vmark}
Given the continuities described in \ref{fig:math:graphic:retraction:map}, 
we illustrate a minimal Q that will generate the most minimal visualizations associated with those continuities: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \caption{The data is discrete points (time, precipitation). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase\ via $\vindex^*$. The pulled back visual secion $\vchannel$ is composited with the assembly function $\vmarkd \circ  \vchannel=\gsection$ to produce the instructions to make the graphic \gsection. The graphic section fills in the pixels in the screen via lookup on \gbase.}
    \label{fig:math:artist:q:scatter}
\end{figure}
The scatter plot in \autoref{fig:math:artist:q:scatter} has a constant size and color $\gsection_{RGB} = (0,0,0)$ that are defined as part of the point assembly function.

>>>>

\section{Topological Equivariant Artist Model}
\label{sec:math}
To guide the implementation of structure preserving visualization components, we develop a mathematical formalism of visualization that specifies how these components preserve \textit{continuity} and \textit{equivariance}. Inspired by the analogous classes in Matplotlib\cite{hunterArchitectureOpenSource}, we call the transformation from data space to graphic space that these building block components implement the \textit{artist}.
\begin{equation}
    \label{eq:artist}
    \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
\end{equation}
The \textit{artist} $\mathscr{\vartist}$ is a map from data $\mathscr{\dtotal}$ to graphic $\mathscr{\gtotal}$ fiber bundles. To explain how the \textit{artist} is a structure preserving map from data to graphic, we first model data (\autoref{sec:math:data}) and graphics (\autoref{sec:math:graphic}) as topological structures that encapsulate component types and continuity. We then discuss the functional maps from graphic to data (\autoref{sec:math:graphic:base}), data components to visual components (\autoref{sec:math:artist:nu}), and visual components into graphic (\autoref{sec:math:artist:q}) that make up the artist.

@@

A monoid \cite{Monoid2021} $\monoid$ is a set with a binary operation $\ast:\monoid \times \monoid\rightarrow \monoid$ that satisfies the axioms:
\begin{align*}
    \textbf{associativity}\; & \text{for all} a, b, c \in \monoid \, (a \bullet b) \bullet c = a \bullet (b \bullet c)\\
    \textbf{identity}\; & \text{for all } a\in \monoid,\,  e\bullet a = a 
\end{align*} 
As defined on a component of \dfiber, a left monoid action \cite{SemigroupAction2021,nlabauthorsAction2021} of $\monoid_i$ is a set $\dfiber_i$ with an action $\bullet: \monoid\times \dfiber_i \rightarrow \dfiber_i$ with the properties:
\begin{align*}
    \textbf{associativity}\;& \text{for all } f,g \in \monoid_i \text{ and } x\in \dfiber_i,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
    \textbf{identity}\;& \text{for all } x\in \dfiber_i, e\in \monoid_i,\,  e\bullet x = x 
\end{align*}
We conjecture that given these properties, if there is a partial ordering on both sides of the action then the action is equivariant and monotonic.

\begin{figure}[H]
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=\textwidth]{figures/math/monoid_total.png}
        \caption{}
        \label{fig:math:data:monoid:total}
    \end{subfigure}
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=1\textwidth]{figures/math/monoid_partial.png}
        \caption{}
        \label{fig:math:data:monoid:partial}
    \end{subfigure}
    \caption{The action $\bullet$ in \autoref{fig:math:data:monoid:total} is the arrows from the partial order diagram of weather states on the left to the diagram of weather states on the right. The action maps the weather states to themselves, thereby preserving the ordering defined by the monoid $\ast$ on both sides of the action. We conjecture that the action in \autoref{fig:math:data:monoid:partial} is monotonic when the partial ordering of the weather states is the same as the partial order of the elements they are mapped to. Given $sun \geq clouds\geq rain,\, lightining$ on the right, we conjecture that the action  $sun,\, clouds \rightarrow clouds$, and $rain,\, lightining \rightarrow rain$ is structure preserving such that the relative ordering of elements is the same as the elements they are mapped to.}
    \label{fig:math:data:monoid}
\end{figure}

@@

As with the fiber \dfiber\, the total monoid space \monoid\ is the cartesian product
\begin{equation}
    \label{eq:math:data:monoid:decompose}
\monoid= \monoid_{0} \times \ldots \times \monoid_{i}\times \ldots \times\ldots \monoid_{n}
\end{equation}
of each monoid \(\monoid_{i}\) on \(\dfiber_{i}\).  The monoid is added to the specification of the fiber \((\fname_i,\, \ftype_i,\, \fttype\, \monoid_i)\)

\subsubsection{Continuity of the Data} 
\label{sec:math:data:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space \dbase\ encodes the continuity of the data space. The table of discrete weather station records has discrete continuity such that each record maps to a single point (a). A gaussian has a value at all points along the interval x is sampled from and therefore has a 1D continuity (b). The globe has a value at all points (latitude, longitude) on the globe and therefore has 2D continuity (c). }
    \label{fig:math:data:base:types}
\end{figure}

The base space \dbase\ provides a way to explicitly encode the continuity of the data, as described in \autoref{sec:intro:data:continuity}. This explicit topology is a concise way of distinguishing between visualizations that appear identical but assume different continuity, for example heat maps and images. The base space \dbase\ acts as an indexing space, as emphasized by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, \dbase\ can have any number of dimensions and can be continuous or discrete. Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $\dbasepoint \in \dbase$ has a corresponding fiber $\dfiber_k$\cite{QuotientSpaceTopology2020}.

@@

In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping $\pi: \dtotal\rightarrow \dbase$ is the closed interval $\left[0,1\right]$. While the total space \dtotal\ may have components in \dfiber\ that describe any given point $\dbasepoint \in \dbase$, such as \textit{time, latitude, longitude}, these labels are indexed into from \dbase\ the same as any other components. In contrast to the structural \textit{keys} with associated \textit{values} proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, our model treats keys \dbasepoint\ as a pure reference to topology. Decoupling the keys from their semantics allows the components identifying the keys to be altered, which provides for coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example total rainfall is dependent on time of day and how much rain has already fallen, and changing the coordinate system or time resolution should have no effect on how the records are connected to each other.

@@

\subsubsection{Data Values}
\label{sec:math:data:section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section can be decomposed as 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber (\autoref{eq:math:data:base:decompose}, \autoref{eq:math:data:fiber:decompose}), we can decompose \dsection\ as 
\begin{equation}
\label{eq:math:data:section:decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \caption{ Fiber (time, precipitation) with a 1D continuous \dbase\  defined on an interval $\left[0,n\right]$. The sections $\dsection^{(1)}$ and $\dsection^{(2)}$ are constrained such that the time variable must be monotonic, which means each section is a time series of precipitation values. They are included in the global set of sections  $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$}
    \label{fig:math:data:sections}
\end{figure}

In \autoref{fig:math:data:sections}, the fiber is the same encoding of $(time, \, precipitation)$ illustrated in \autoref{fig:math:data:fiber}, and the base space is the interval \dbase\ shown in \autoref{fig:math:data:base}. The section $\dsection^{(1)}$ is a function that for a point \dbasepoint\ returns a record in the fiber \dfiber. The section applied to a set of points in \dbase\ resolves to a series of monotonically increasing in time records of \textit{(time, precipitation)} values. Section $\dsection^{(2)}$ returns a different time series of \textit{(time, precipitation)} values. Both sections are included in the global set of sections $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$.

@@

\subsubsection{Applications}
Using fiber bundles as our data abstraction allows the model to be applicable to a wide variety of fields and use cases. For example, the section can be any instance of a numpy array\cite{harris2020array} that stores an image, such as an image where the \dbase\ is a 2D continuous plane and the \dfiber\ is $(\reals^{3}, \reals, \reals)$. In this fiber, the $\reals^3$ components encode color, and the other two components are the x and y positions of the sampled data in the image. The continuity of the image is implicitly encoded in the array as the index, so the position components encode the resolution. Instead of an image, the numpy array could also store a 2D discrete table. The fiber may not change, but the \dbase\ would now be 0D discrete points. These different choices in topology indicate, for example, what sorts of interpolation would be appropriate when visualizing the data. Labeled containers can also be described in this framework because of the schema like structure of the fiber. One such example is a pandas series which stores a labeled list, another is a  dataframe\cite{jeff_reback_2020_3715232} which has the structure of a relational table. A series could store the values of $\dsection^{(1)}$ and a second series could be  $\dsection^{(2)}$, while a dataframe would have multiple components and each data frame would be a unique section \dsection. The ability to encode complexity in continuity and components is particularly beneficial when working with  N dimensional labeled data containers. For example, an xarray\cite{hoyer2017xarray} data cube that stores precipitation would be a section of a fiber bundle with a \dbase\ that is a continuous volume and components \textit{(time, latitude, longitude, precipitation)}. This section does not need to resolve to values immediately and instead can be an instance of a distributed data container, such as a dask array \cite{rocklinDaskParallelComputation2015}. Our model provides a common formalism for describing widely used data containers without sacrificing the semantic structure embedded in each container.

@@

\subsubsection{Graphic}
\label{sec:math:graphic:section}
The section $\gsection: \gbase \rightarrow \gtotal$ is the graphic in an idealized prerender space and also acts as a specification for rendering the graphic to target display format. To demonstrate the role of \gsection\, it is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{x,y}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. The color is the result of the integration over  $\gbase_p$.

@@

\subsubsection{Visual Encoders}
\label{sec:math:artist:nu}
We propose that the map from data components to graphic components $\vchannel: \dsection \mapsto \vsection$ is a monoid \textit{equivariant} map. By specifying this constraints, we can guarantee that the stage of the artist that transforms data components into graphic representations is equivariant. These constraints then guide the implementation of reusable component transformers \vchannel\ that are composed when generating the graphic. We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. Therefore, a validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
such that applying equivariant monoid actions to $\dtotal_{i}$ and $\vtotal_{i}$ preserves the map $\vchannel_{i}: \dtotal_{i}\rightarrow \vtotal_{i}$. In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\label{eq:math:artist:nu:equivariance}
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, 
without a loss of generality we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function. We can make this assumption because we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can then compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$, which lets us incorporate $\varphi$ into the action $\bullet$ such that $\varphi$ does not need to be explicitly defined in the constraints.

@@

\subsubsection{Visualization Assembly}
Having described the maps to components in \autoref{sec:math:artist:nu}, we now specify the assembly function \vmarkd\ that composites components in \vtotal\ into a graphic in \gtotal. Since the component transforms \vchannel\ are equivariant, the equivariance constraints carry through to \vmarkd. We specify these constraints to guide the implementation of library components responsible for generating graphics.

@@

\subsubsection{Assembly \vmark}
Given the continuities described in \ref{fig:math:graphic:retraction:map}, 
we illustrate a minimal Q that will generate the most minimal visualizations associated with those continuities: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \caption{The data is discrete points (time, precipitation). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase\ via $\vindex^*$. The pulled back visual section $\vchannel$ is composited with the assembly function $\vmarkd \circ  \vchannel=\gsection$ to produce the instructions to make the graphic \gsection. The graphic section fills in the pixels in the screen via lookup on \gbase.}
    \label{fig:math:artist:q:scatter}
\end{figure}
The scatter plot in \autoref{fig:math:artist:q:scatter} has a constant size and color $\gsection_{RGB} = (0,0,0)$ that are defined as part of the point assembly function.
